---
title: "Introduction to opencgaR"
author:
  - Marta Bleda
  - Mohammed Omar Elsiddieg Abdallah
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('opencgaR')`"
vignette: >
  %\VignetteIndexEntry{"Introduction to opencgaR"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
output:
   BiocStyle::html_document
---
# Introduction
OpenCGA is an open-source project that aims to provide a **Big Data storage engine and analysis framework for genomic scale data analysis** of hundreds of terabytes or even petabytes. For users, its main features include uploading and downloading files to a repository, storing their information in a generic way (non-dependant of the original file-format) and retrieving this information efficiently. For developers, it will be a platform for supporting the most used bioinformatics file formats and accelerating the development of visualization and analysis applications.

The OpencgaR client provides a user-friendly interface to work with OpenCGA REST Web Services through R and has been implemented following the Bioconductor guidelines for package development which promote high-quality, well documented and interoperable software. The source code of the R package can be found in [GitHub](https://github.com/opencb/opencga/tree/develop/opencga-client/src/main/R). For more information see the documentation [website](http://docs.opencb.org/display/opencga/Overview).

# Installation and configuration
Please visit http://docs.opencb.org/display/opencga/Installation+Guide for detailed information on how to install and configure OpenCGA.

# Users login and authentication
A set of methods have been implemented to deal with the connectivity and login to the REST host. Connection to the host is done in two steps using the functions *__initOpencgaR__* and *__opencgaLogin__* for defining the connection details and loging in, respectively.

The *__initOpencgaR__* function accepts either host and version information or a configuration file (as `list()` or in [YAML or JSON format](http://docs.opencb.org/display/opencga/client-configuration.yml)). The *__opencgaLogin__* function stablishes the connection with the host. It requires an object opencgaR (created using *initOpencgaR*) and the login details: user and password. User and password can be introduced interactively through a popup window using `interactive=TRUE`, to avoid typing user credentials in the R script.

The code below shows the different ways to initialise the OpenCGA connection with the REST server.

```{r, eval=FALSE}
# Initialise connection specifying host and REST version
con <- initOpencgaR(host = "http://localhost:8080/opencga/", version = "v1")

# Initialise connection using configuration (list)
conf <- list(version="v1", rest=list(host="http://localhost:8080/opencga/"))
con <- initOpencgaR(opencgaConfig=conf)

# Initialise connection using configuration (YAML or JSON format)
conf <- "/path/to/conf/client-configuration.yml"
con <- initOpencgaR(opencgaConfig=conf)
```

Once the connection has been initialised users can login specifying their OpenCGA user ID and password.

```{r, eval=FALSE}
# Log in
con <- opencgaLogin(opencga = con, userid = "user", passwd = "pass")
```




# Navigation through users projects, studies and files
# Querying variant data
# Integration with the Bioconductor ecosystem
